---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
    toc: true
---

##1. Introduction

This project will focus on attributes make a movie popular as well as how much audiences and critics like movies and numerous other variables about the movies. The dataset was randomly collected from Rotten Tomatoes and IMDB in the US. 

***

## 2. Set up

### 2.1 Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(statsr)
library(BAS)
library(MASS)
library(grid)
library(GGally)
library(gridExtra)
```

### 2.2 Load data

```{r load-data}
load("movies.Rdata")
```

**Scope of Inference**: This dataset includes information from Rotten Tomatoes and IMDB for a random sample of movies. Thus, the study is obsevational and only shows associational relationship.

The movies data from IMDB was used for the analysis at hand. Some of variables are in the original dataset provided, and others are new variables. This will need be to construct in the data manipulation section. 

  - feature_film: 'yes' if title_type is Feature Film, 'no'         otherwise
  - drama: "yes" if genre is Drama, "no" otherwise
  - mpaa_rating_R: "yes" if mpaa_rating is R, "no" otherwise
  - runtime
  - thtr_rel_year 
  - oscar_season: "yes" if movie is released in November,           October, or December (based on thtr_rel_month), "no"            otherwise
  - summer_season: "yes" if movie is released in May, June,         July, or August (based on thtr_rel_month), "no" otherwise
  - imdb_rating
  - imdb_num_votes
  - critics_score
  - best_pic_nom 
  - best_pic_win
  - best_actor_win 
  - best_actress_win
  - best_dir_win
  - top200_box
  
```{r}
# evaluate the size of the dataset
dim(movies)
```

```{r}
# types and summary of each variable 
str(movies)
summary(movies)
```

We see that there are some NA values for some variables. In this project, we will assume that the missing NA is not important and will not affect the result of the analysis so much that we should concern. So, we can eliminate all NA values

```{r}
movies_no_na <- na.omit(movies)
dim(movies_no_na)
```

The data has 619 observations and 37 variables. 


### 2.3 Data manipulation

**`feature_film`: "yes" if title_type is Feature Film, "no" otherwise**

```{r}
movies_no_na <- mutate(movies_no_na, feature_film = ifelse(title_type == "Feature Film", "Yes", "No"))
movies_no_na$feature_film <- as.factor(movies_no_na$feature_film)
summary(movies_no_na$feature_film)
```


**`drama`: "yes" if genre is Drama, "no" otherwise**

```{r}
movies_no_na <- mutate(movies_no_na, drama = ifelse(genre == "Drama", "Yes", "No"))
movies_no_na$drama <- as.factor(movies_no_na$drama)
summary(movies_no_na$drama)
```


**`mpaa_rating_R`: "yes" if mpaa_rating is R, "no" otherwise**

```{r}
movies_no_na <- mutate(movies_no_na, mpaa_rating_R = ifelse(mpaa_rating == "R", "Yes", "No"))
movies_no_na$mpaa_rating_R <- as.factor(movies_no_na$mpaa_rating_R)
summary(movies_no_na$mpaa_rating_R)
```


**`oscar_season`: "yes" if movie is released in November, October, or December (based on thtr_rel_month), "no" otherwise**

```{r}
movies_no_na <- mutate(movies_no_na, oscar_season = ifelse(thtr_rel_month %in% c(10,11,12), "Yes", "No"))
movies_no_na$oscar_season <- as.factor(movies_no_na$oscar_season)
summary(movies_no_na$oscar_season)
```


**`summer_season`: "yes" if movie is released in May, June, July, or August (based on thtr_rel_month), "no" otherwise**

```{r}
movies_no_na <- mutate(movies_no_na, summer_season = ifelse(thtr_rel_month %in% c(5,6,7,8), "Yes", "No"))
movies_no_na$summer_season <- as.factor(movies_no_na$summer_season)
summary(movies_no_na$summer_season)
```


```{r}
df <- movies_no_na[c("feature_film","drama","mpaa_rating_R","oscar_season","summer_season","audience_score")]

summary(df)
```

* * *

## 3 Exploratory Data Analysis

**First, we will overview the association of the audience score and feature film**

```{r}
summary(df$audience_score)
IQR(df$audience_score)
mean(df$audience_score)
```

The IQR of the audience_score is 34 (1st Qu - 46 and 3rd Qu. 80), while the mean is around 62.2, the median is 65

```{r}
ggplot(df, aes(x = audience_score, y = ..density..)) +
  geom_histogram(bins = 40, fill = 'blue', colour = 'black') + 
  geom_density(size = 1, colour = 'brown') 
```

The audience_score shows a slight left skewed structure. 

We will create a boxplot for each feature and compare them with the audience_score. We will also analyse the variability of the new features by comparing them to each other.


```{r fig.width = 12, fig.height = 10}
p1 <- ggplot(df, aes(y=audience_score, x=feature_film, fill = feature_film)) +
  geom_boxplot() +
  ggtitle("Relationship between Audience Score and Feature Films") +
  xlab("Feature Films") +
  ylab("Audience Score") + 
  scale_fill_brewer(name = "feature film")


p2 <- ggplot(df, aes(x = summer_season, y = audience_score, fill = summer_season)) + 
  geom_boxplot() + 
  ggtitle('Audience score vs summer season') + 
  xlab('summer season') + 
  ylab('Audience Score') + scale_fill_brewer(name = "summer season")


p3 <- ggplot(df, aes(x = oscar_season, y = audience_score, fill = oscar_season)) + 
  geom_boxplot() + 
  ggtitle('Audience score vs oscar_season') +
  xlab('oscar_season') + 
  ylab('Audience Score') + 
  scale_fill_brewer(name = "oscar_season")


p4 <- ggplot(df, aes(x = drama, y = audience_score, fill = drama)) + 
  geom_boxplot() +
  ggtitle('Audience score vs drama') + 
  xlab('drama') + 
  ylab('Audience Score') + 
  scale_fill_brewer(name = "drama")


p5 <- ggplot(df, aes(x = mpaa_rating_R, y = audience_score, fill = mpaa_rating_R)) + 
  geom_boxplot() + 
  ggtitle('Audience score vs mpaa_rating_R') +
  xlab('mpaa_rating_R') + 
  ylab('Audience Score') + 
  scale_fill_brewer(name = "mpaa_rating_R")

# arrange the previously created plots 
grid.arrange(p1, p2, p3, p4, p5, ncol = 2)
```

From above plots, we can see that:
  
  - Audience score for no feature film is higher than featur        film. 
  - Drama movies have higher avergae audience score. 
  - Audience score for no mpaa_rating_R is slightly higher than     mpaa_rating_R. 
  - Audience score for movies released in Nov, Oct or Dec has       higher score than movies released in other months. 

Finally we will have a look at the other features used in the dataset

```{r fig.width=12, fig.height=10}

pf1 <- ggplot(df, aes(audience_score, fill = feature_film)) + 
  geom_density() + 
  ggtitle("Audience score vs. feature_film") + 
  labs(x = "feature film", y = "Density")

    
pf2 <- ggplot(df, aes(audience_score, fill = drama)) +
  geom_density () + 
  labs(title = "Audience score vs. drama") + 
  labs(x = "drama", y = "Density")


pf3 <- ggplot(movies_no_na, aes(audience_score, fill = top200_box))+ 
  geom_density () + 
  labs(title = "Audience score vs. top200_box") + 
  labs(x = "top200 box", y = "Density")

    
pf4 <- ggplot(df, aes(audience_score, fill = oscar_season)) + 
  geom_density() + 
  labs(title = "Audience score vs. oscar_season") +
  labs(x = "oscar season", y = "Density")


pf5 <- ggplot(df, aes(audience_score, fill = summer_season))+ 
  geom_density () + 
  labs(title = "Audience score vs. summer_season") + 
  labs(x = "summer season", y = "Density")

    
pf6 <- ggplot(movies_no_na, aes(audience_score, fill = best_pic_nom))+ 
  geom_density () + 
  labs(title = "Audience score vs. best_pic_nom") + 
  labs(x = "best pic nom", y = "Density")


pf7 <- ggplot(movies_no_na, aes(audience_score, fill = best_pic_win)) + 
  geom_density() + 
  labs(title = "Audience score vs. best pic win") + 
  labs(x = "best pic win", y = "Density")

    
pf8 <- ggplot(movies_no_na, aes(audience_score, fill = best_actor_win))+ 
  geom_density () + 
  labs(title = "Audience score vs. best_actor_win") + 
  labs(x = "best actor win", y = "Density")

  
pf9 <- ggplot(movies_no_na, aes(audience_score, fill = best_dir_win))+ 
  geom_density () + 
  labs(title = "Audience score vs. best_dir_win") + 
  labs(x = "best dir win", y = "Density")

    
pf10 <- ggplot(movies_no_na, aes(audience_score, fill = best_actress_win))+ 
  geom_density () + 
  labs(title = "Audience score vs. best_actress_win") + 
  labs(x = "best actress win", y = "Density")

   
grid.arrange(pf1, pf2, pf3, pf4, pf5, pf6, pf7, pf8, pf9, pf10, ncol = 2)
```


##4. Hypothesis testing

we will use the bayes_inference function, which will allow us to construct credible intervals perform a hypothesis test and calculate Bayes factors for a variety of different circumstances. The main goal is to investigate if the newly created features(feature_film, drama, mpaa_rating_R, oscar_season and summer_season) influence the audience_score. 

```{r}
bayes_inference(y = audience_score, x = feature_film, data = df, statistic = 'mean', type = 'ht', null = 0, alternative = 'twosided', prior = "JZS", rscale = 1, method = "theoretical")

bayes_inference(y = audience_score, x = drama, data = df, statistic = 'mean', type = 'ht', null = 0, alternative = 'twosided', prior = "JZS", rscale = 1, method = "theoretical")

bayes_inference(y = audience_score, x = mpaa_rating_R, data = df, statistic = 'mean', type = 'ht', null = 0, alternative = 'twosided', prior = "JZS", rscale = 1, method = "theoretical")

bayes_inference(y = audience_score, x = oscar_season, data = df, statistic = 'mean', type = 'ht', null = 0, alternative = 'twosided', prior = "JZS", rscale = 1, method = "theoretical")

bayes_inference(y = audience_score, x = summer_season, data = df, statistic = 'mean', type = 'ht', null = 0, alternative = 'twosided', prior = "JZS", rscale = 1, method = "theoretical")
```


**Summary**

feature_film: BF[H2:H1] = 14.6e+9 --> Evidence against H1 is Very Strong

drama: BF[H2:H1] = 50.77 --> Evidence against H1 is positive    

mpaa_rating_R: BF[H1:H2] = 15.354 --> Evidence against H2 is Positive

oscar_season: BF[H1:H2] = 6.2077 --> Evidence against H2 is Positive

summer_season: BF[H1:H2] = 14.0392  --> Evidence against H2 is Positive


For feature_film, there is strong evidence against H1, which means that there is a significant difference in mean audience_score for feature- and non-feature films


* * *

## 5. Modeling

Based on the research question, we will just focus on the relationship between audience_score and variables mentioned in part 2. In this section, we will use backwards elimination to pick significant predictors and first, we will start with full model.


```{r}
data.model <- movies_no_na[c("feature_film","drama","runtime","mpaa_rating_R","thtr_rel_year","oscar_season","summer_season","imdb_rating","imdb_num_votes","critics_score","best_pic_nom","best_pic_win","best_actor_win","best_actress_win","best_dir_win","top200_box","audience_score")]

str(data.model)
```

```{r}
lm1 <- lm(audience_score ~ ., data = data.model)
score_step <- stepAIC(lm1, trace = FALSE)
score_step$anova
```

We do not see mpaa_rating_R in this model but summer_season, oscar_season are still included in this model. We will use Bayesian Model Average (BMA) next

```{r}
bma_audience_score <- bas.lm(audience_score ~., data = data.model, prior = "BIC", modelprior = uniform())
bma_audience_score
```

```{r}
summary(bma_audience_score)

```

```{r fig.width= 12, fig.height=10}
image(bma_audience_score, rotate = FALSE)

```


We can see from the model rank that there are three variables that have high posterior odds which are runtime, imdb_rating, critic_score.

```{r}
coef_bma_audience_score <- coef(bma_audience_score)
plot(coef_bma_audience_score)
```


Now we can build up a model for prediction:

```{r}
finalmodel <- data.model[c("runtime","imdb_rating","critics_score","audience_score")]

bma_finalmodel <- bas.lm(audience_score ~., data = finalmodel, prior = "ZS-null", method = "MCMC", modelprior = uniform())

summary(bma_finalmodel)
```



* * *

## 6. Prediction

Build test data cases for the movie "Black Panther (2018)" using the data gathered from IMDB (imdb_rating = 7.4) and rotten tomatoes website (audience_score = 79) and storing the data in the variable named blackpanther (test data case) using the following code

```{r}
blackpanther <- data.frame(feature_film="yes",drama="no",runtime=135,mpaa_rating_R="no",thtr_rel_year=2018,oscar_season="no",summer_season="no",imdb_rating=7.4,imdb_num_votes=443501,critics_score=97,best_pic_nom="no",best_pic_win="no",best_actor_win="no",best_actress_win="no",best_dir_win="no",top200_box="yes",audience_score=79)

data.predict <- rbind(data.model, blackpanther)

blackpanther <- tail(blackpanther, 1)

str(blackpanther)

```

We will predict the audience_score

```{r}
audience_score_prediction <-predict(bma_finalmodel, newdata=blackpanther, estimator="BMA", se.fit=TRUE, interval="predict", level = 0.95)
audience_score_prediction$Ybma
```


The prediction is lower than the actual audience_score.

* * *

## 7. Conclusion

The project uses data from movies dataset to determine if there is any association between audience_score and other variables. Doing exploratory data analysis and modeling help us to know which variables are significant predictors. Yet, together with Bayes model, we can see different model and can pick the model that have the highest prediction.